---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# WSSmod

<!-- badges: start -->
<!-- badges: end -->

WSSmod provides tools to construct Weighted Sum of Scores (WSS) module-level
summaries from proteomic data. Proteins are grouped into STRING network
modules, sign-flipped to align with the direction of core biomarker
association, and combined into a single module score per sample using a
weighted average based on network connectivity (alpha centrality) and
biomarker association strength.

This package accompanies the Longo Lab plasma biomarker proteomics work.

## Installation

You can install the development version of WSSmod from GitHub with:

``` r
# install.packages("pak")
pak::pak("Longo-Lab/WSSmod")
```

## Example

```{r example}
library(WSSmod)

expr_matrix <- matrix(
  c(1, 2, 3, 4, 5, 6, 7, 8),
  nrow = 2,
  dimnames = list(c("sample1", "sample2"), c("geneA", "geneB", "geneC", "geneD"))
)

result <- data.table::data.table(
  module = c("mod1", "mod1", "mod1", "mod2"),
  symbol = c("geneA", "geneB", "geneC", "geneD"),
  mean_beta = c(1, -1, 1, 1),
  mean_alpha_scaled = c(1, 1, 2, 1)
)

WSSmod::calculate_weighted_module_scores(expr_matrix, result)
```
